---
title: "Untitled"
author: "Gabe Runte"
date: "9/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)
library(here)
library(janitor)
library(gridExtra)
library(RColorBrewer)

```

First, I will pull in all of the sequences that I want to assign taxonomy to. Workflow is as follows:
1. Pull all of the consensus sequences
2. Combine with list of sequences not assigned to a contig
3. Map these onto the total list of sequences, allowing one to then remove samples with short sequences or with <2% high quality bp reads
4. Create final list of sequences to be assigned taxonomy
```{r, message = F}
consensus = read_csv(here("data/sanger_data/morpho_nov162023", "consensus.csv")) %>% 
  clean_names() %>% 
  dplyr::select(name, sequence) %>% 
  rename(sequence_list_name = name)

contig_assign = read_csv(here("data/sanger_data/morpho_nov162023", "contigs.csv")) %>% 
  clean_names() %>% 
  rename(contig_sequence = sequence) %>% 
  left_join(consensus)

metadata = read_csv(here("data/sanger_data/morpho_nov162023", "all_seqs.csv")) %>% 
  clean_names() %>% 
  rename(original_seq = sequence) %>% 
  select(!sequence_list_name) %>% 
  left_join(contig_assign) %>% 
  mutate(sequence = case_when(
    is.na(sequence_list_name) ~ original_seq,
    !is.na(sequence_list_name) ~ sequence
  )) %>% 
  select(!original_seq)%>% 
  select(!contig_sequence) %>% 
  filter(!is.na(sequence)) %>% 
  filter(post_trim_length>75)
```



```{r, eval = FALSE}
unite.ref <- "/Users/Gabe/Desktop/phd/ghecto/unite_db/sh_general_release_s_10.05.2021/sh_general_release_dynamic_s_10.05.2021.fasta"
unique_sequences = unique(metadata$sequence)
tiptaxa = dada2::assignTaxonomy(unique_sequences, unite.ref, multithread = TRUE, tryRC = TRUE)


unite_taxa = tibble(as.data.frame(tiptaxa)) %>% 
  mutate(sequence = rownames(tiptaxa))
write_csv(unite_taxa, here("data/sanger_data/morpho_nov162023", "unite_output.csv"))
```


```{r}
unite_taxa = read_csv(here("data/sanger_data/morpho_nov162023", "unite_output.csv")) 
sample_sheet = read_csv(here("data/harvest", "harvest_tipseq_printout.csv")) %>% 
  mutate(sampleset = "m5x") %>% 
  bind_rows(read_csv(here("data/harvest", "harvest_tipseq_printout2.csv")) %>% 
  mutate(sampleset = "m2x")) %>% 
  select(-plate, -column, -row)
seedlings = read_csv(here("data/analysis","size_master.csv" )) %>% 
  select(seedling_id, seedling_sp, soil_type, neighb_sp, neighb_st) %>% 
  distinct() %>% 
  filter(!is.na(neighb_sp))

tip_seqs = metadata %>% 
  left_join(unite_taxa) %>% 
  clean_names() %>% 
  separate(name, into = c("project", "sampleset", "comb_column", "comb_row")) %>% 
  mutate(comb_column = as.numeric(comb_column)) %>% 
  select(-project) %>% 
  left_join(sample_sheet) %>% 
  left_join(seedlings) %>% 
  filter(!is.na(seedling_sp))%>% 
  mutate(tree_div = case_when(
    seedling_sp == neighb_sp ~1,
    .default = 2
  )) %>% 
  filter(!is.na(phylum)) %>% 
  mutate(taxonomy = paste(kingdom, phylum, class, order, family, genus, species, sep = ";")) %>% 
  mutate(sample = paste(sampleset, comb_column, comb_row, sep = "_"))



fung_df = tip_seqs %>% 
  select(sample, taxonomy)

write_csv(fung_df, here("data/sanger_data/morpho_nov162023", "taxa_funguild.csv"))
#open file with xcel or other spreadsheet software then convert to .txt file at same file path (see below)
here("data/sanger_data/morpho_nov162023", "taxa_funguild.txt")
```

```{bash, engine.opts='-l', eval = F}

python /Users/Gabe/Desktop/phd/FUNGuild-master/Guilds_v1.1.py -otu /Users/Gabe/Desktop/phd/ghecto/data/sanger_data/morpho_nov162023/taxa_funguild.txt
```

```{r}

guilds = read_csv(here("data/sanger_data/morpho_nov162023", "taxa_funguild.guilds.csv")) %>% 
  clean_names()

tip_all = tip_seqs %>% 
  left_join(guilds) %>%
  filter(str_detect(guild, "Ectomycorrhizal")) %>% 
  mutate(n_splong = case_when(
    neighb_sp == "Q" ~ "Oak neighbor",
    neighb_sp == "P" ~ "Pine neighbor"
  ))

ggplot(tip_all, aes(x = morphotype, fill = genus))+
  geom_bar()
ggplot(tip_seqs, aes(x = morphotype, fill = genus))+
  geom_bar()

ggplot(tip_all, aes(x = morphotype, fill = genus))+
  geom_bar()+facet_grid(rows = vars(seedling_sp), cols = vars(n_splong))


ggplot(tip_seqs, aes(x = morphotype, fill = genus))+
  geom_bar()+facet_grid(rows = vars(seedling_sp), cols = vars(neighb_sp))
```

Planning out which morphotypes to target:
A - Pustularia (DONE)
B - Tuber (DONE)
C - Unknown (UNKNOWN) - might not be doable 
D - Unknown (MIXED) - what to do here... assign relatively? 
E - Unknown (MIXED) - what to do here... assign relatively? 
F - Pustularia (DONE)
G - Pustularia (DONE)
H - Possibly melanogaster (10 more)
o - Pustularia, although rarely successfully ectomycorrhizal
W - Rhizopogon (5 more)
X - Rhizopogon (DONE)
Y - Non ecto - off the list
Z - Pustularia (5 more)

Gameplan will be to sequence 5 more of each of the morphotypes we are confident on. The rest, I will pursue in sets of 10 sequences. This includes getting a rate of amplification success for the 'O' morphotype. 



## 
Now lets combine this dataset with our knowledge of how many samples failed at PCR vs poor Seq quality, etc
```{r}
all_attempts  = read_csv(here("data/harvest", "harvest_tipseq_printout.csv")) %>% 
  mutate(sampleset = "m5x") %>% 
  bind_rows(read_csv(here("data/harvest", "harvest_tipseq_printout2.csv")) %>% 
  mutate(sampleset = "m2x")) 

all_seqs = read_csv(here("data/sanger_data/morpho_nov162023", "all_untrimmed.csv")) %>% 
  clean_names()%>% 
  separate(name, into = c("project", "sampleset", "comb_column", "comb_row")) %>% 
  mutate(comb_column = as.numeric(comb_column)) %>% 
  select(-project) %>% 
  left_join(all_attempts) %>% 
  left_join(seedlings)


total_df = all_attempts %>% 
  select(seedling_id, plate, column, row, morphotype) %>% 
  left_join(all_seqs) %>% 
  left_join(tip_all %>% 
              select(1:3, genus, species, guild, n_splong))  %>% 
  mutate(genus = str_sub(genus, 4))%>% 
  mutate(if_genus = case_when(
    is.na(sequence)~ "PCR Fail",
    is.na(genus)~ "Bad Seq",
    .default = genus
  )) %>% 
  select(-seedling_sp, -soil_type, -neighb_sp, -neighb_st) %>% 
  left_join(seedlings) %>% 
  filter(!is.na(seedling_sp)) %>% 
  mutate(if_genus = fct_relevel(if_genus, "Bad Seq", "PCR Fail", "Melanogaster", "Pustularia", "Rhizopogon", "Tuber", "Wilcoxina" )) %>% 
  mutate(sp_long = case_when(
    seedling_sp =="Q" ~"Oak seedling",
    seedling_sp =="P" ~"Bigcone seedling"
  ))%>% 
  mutate(nsp_long = case_when(
    neighb_sp =="Q" ~"Oak neighbor",
    neighb_sp =="P" ~"Bigcone neighbor"
  ))

ggplot(total_df, aes(x = morphotype, fill = if_genus))+
  geom_bar()+facet_grid(rows = vars(seedling_sp), cols = vars(neighb_sp))+
  scale_fill_manual(breaks = c("Melanogaster", "Pustularia", "Rhizopogon", "Tuber", "Wilcoxina", "Bad Seq", "PCR Fail"),
                     name = "Genus", values = c("#e65247","#e89923", "#f0dc5d", "#4dc44d", "#4287f5", "darkgray", "gray"))+theme_bw()

ggplot(total_df, aes(x = morphotype, fill = if_genus))+
  geom_bar()+facet_grid(rows = vars(seedling_sp))+
  scale_fill_manual(breaks = c("Melanogaster", "Pustularia", "Rhizopogon", "Tuber", "Wilcoxina", "Bad Seq", "PCR Fail"),
                     name = "Genus", values = c("#e65247","#e89923", "#f0dc5d", "#4dc44d", "#4287f5", "darkgray", "gray"))+
  theme_minimal()



ggplot(total_df, aes(x = as.factor(seedling_id), fill = if_genus))+lims(y = c(0,4))+
  geom_bar()+facet_grid(rows = vars(sp_long), cols = vars(nsp_long), scales = "free")+
  scale_fill_manual(breaks = c("Melanogaster", "Pustularia", "Rhizopogon", "Tuber", "Wilcoxina", "Bad Seq", "PCR Fail"),
                     name = "Genus", values = c("#e65247","#e89923", "#f0dc5d", "#4dc44d", "#4287f5", "darkgray", "gray"))+theme_minimal()+labs(y = "Sequence count", x= "Seedling")+ theme(
  axis.text.x = element_blank())
```


FROM HOLLY: Rethink D,E,H,O,W

Rethinking successes:
A - Pustularia (DONE)
B - Tuber (DONE)
C - Unknown (UNKNOWN) - Too few tips to really matter 
D - Unknown (MIXED) - what to do here... assign relatively? 
E - Unknown (DONE) - Have attempted every possible E tip...
F - Pustularia (DONE)
G - Pustularia (DONE)
H - Melanogaster with relatively low success... (UNK)
N - Successfully non-mycorrhizal
o - Pustularia (What is the mycorrhizal rate?)
W - Rhizopogon (What is the mycorrhizal rate?)
X - Rhizopogon (DONE)
Y - Non ecto - (DONE)
Z - Pustularia (5 more)






